Covariate analysis of MSSM reprocessed counts (with CQN normalisation and covariates adjusted)

### 2017-09-05 code/Rmd/MAYO_REPROCESSED.Rmd
Minor modifications and added parallelDuplicateCorrelation.R with lmer

options(xtable.type="html")

@@

# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, 
                         c(postAdjustCovars, primaryVariable, 'Donor_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()

@@

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$Donor_ID)

  # Re-estimate voom weights
  VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, 
                              design=DM1$design, plot=F, 
                              block = COVARIATES$Donor_ID, 
                              correlation = correlation$cor)

@@

  # Residuals after normalisation
  RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                CQN.GENE_EXPRESSION$E)

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates),
                         c(postAdjustCovars, primaryVariable, 'Donor_ID'))

@@

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$Donor_ID)

@@

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS)

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS, block = COVARIATES$Donor_ID)

@@

# Calculate correlation values of random effects
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.SVA.WEIGHTS, block = COVARIATES$Donor_ID)

@@

# Fit linear model using new weights and new design
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.SVA.FIT = lmFit(VOOM.SVA.WEIGHTS)

@@

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t
```

@@

# Calculate correlation values of random effects
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.NET.WEIGHTS, block = COVARIATES$Donor_ID)

@@

# Fit linear model using new weights and new design
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.NET.FIT = lmFit(VOOM.NET.WEIGHTS)

@@

  x[x < (mn-3*std.dev)] = mn-3*std.dev
  x[x > (mn+3*std.dev)] = mn+3*std.dev
  return(x)
}) %>% t
```

@@

# Store differential expression results
rbindlist(all.diff.exp, use.names = T, fill = T, idcol = 'Model') %>%
  write.table(file = 'MAYO_CBE_TCX_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MAYO_CBE_TCX_DiffExpression.tsv', 
                name = 'Differential Expression Results (BrainRegion.Diagnosis)', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r ECOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Surrogate Variables                        |  `r SURR_OBJ$properties$id`      |
|  Residual Expression (for eQTL)             |  `r eEXP_OBJ$properties$id`      |
|  Residual Expression (for network analysis) |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |

>>>>

options(xtable.type="html")
source('../R/parallelDuplicateCorrelation.R')

@@

# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, 
                         c(postAdjustCovars, primaryVariable, 'Donor_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
notEstimable = c()

@@

  # Calculate correlation values of random effects
  VOOM.GENE_EXPRESSION$E = CQN.GENE_EXPRESSION$E
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$Donor_ID,
                                             method = 'lmer')
  # correlation = list(); correlation$cor = 0.457197

  if (!is.nan(correlation$cor)){
    # Re-estimate voom weights
    VOOM.GENE_EXPRESSION = voom(NEW.COUNTS, 
                                design=DM1$design, plot=F, 
                                block = COVARIATES$Donor_ID, 
                                correlation = correlation$cor)

@@

    # Residuals after normalisation
    RESIDUAL.GENE_EXPRESSION = residuals.MArrayLM(VOOM.ADJUSTED.FIT,
                                                  CQN.GENE_EXPRESSION$E)
  } else {
    notEstimable = c(notEstimable, postAdjustCovars[length(postAdjustCovars)])
    postAdjustCovars = postAdjustCovars[1:(length(postAdjustCovars)-1)]
  }

  # Residual covariates to choose from
  residCovars <- setdiff(c(FactorCovariates,ContCovariates),
                         c(postAdjustCovars, primaryVariable, 'Donor_ID', notEstimable))

@@

  # Add postadjusted covariates (if any)
  residCovars = setdiff(residualSigCovars$significantCovars, c(postAdjustCovars, primaryVariable, notEstimable))
  covariatesEffects = residualSigCovars$Effects.significantCovars[residCovars]

@@

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')

@@

# Fit linear model using new weights and new design
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E
FIT = lmFit(VOOM.WEIGHTS$E,
            weights = VOOM.WEIGHTS$weights,
            design = VOOM.WEIGHTS$design,
            block = COVARIATES$Donor_ID, 
            correlation = correlation$cor)

@@

# Estimate voom weights for dispersion control
VOOM.WEIGHTS = voom(NEW.COUNTS, design=DESIGN$design, plot=F)
correlation = parallelDuplicateCorrelation(VOOM.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')

# Calculate correlation values of random effects
VOOM.WEIGHTS$E = CQN.GENE_EXPRESSION$E

@@

# Calculate correlation values of random effects
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.SVA.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')

@@

# Fit linear model using new weights and new design
VOOM.SVA.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.SVA.FIT = lmFit(VOOM.SVA.WEIGHTS$E,
                     weights = VOOM.SVA.WEIGHTS$weights,
                     design = VOOM.SVA.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)

@@

  x[x < (mn-3*std.dev)] = NA
  x[x > (mn+3*std.dev)] = NA
  return(x)
}) %>% t

all.fit = c(all.fit, list(SVA = VOOM.SVA.FIT))
```

@@

# Calculate correlation values of random effects
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
correlation = parallelDuplicateCorrelation(VOOM.NET.WEIGHTS,
                                           block = COVARIATES$Donor_ID,
                                           method = 'lmer')

@@

# Fit linear model using new weights and new design
VOOM.NET.WEIGHTS$E = CQN.GENE_EXPRESSION$E
VOOM.NET.FIT = lmFit(VOOM.NET.WEIGHTS$E,
                     weights = VOOM.NET.WEIGHTS$weights,
                     design = VOOM.NET.WEIGHTS$design,
                     block = COVARIATES$Donor_ID, 
                     correlation = correlation$cor)

@@

  x[x < (mn-3*std.dev)] = mn-3*std.dev
  x[x > (mn+3*std.dev)] = mn+3*std.dev
  return(x)
}) %>% t

all.fit = c(all.fit, list(NET = VOOM.NET.FIT))
```

@@

# Store differential expression results
rbindlist(all.diff.exp, use.names = T, fill = T, idcol = 'Model') %>%
  write.table(file = 'MAYO_CBE_TCX_DiffExpression.tsv', sep = '\t', row.names=F, quote=F)
DEXP_OBJ = File('MAYO_CBE_TCX_DiffExpression.tsv', 
                name = 'Differential Expression Results (BrainRegion.Diagnosis)', 
                parentId = CODE$properties$id)
annotations(DEXP_OBJ) = annotations(CODE)
annotations(DEXP_OBJ)$dataSubType = 'diffExp'
DEXP_OBJ = synStore(DEXP_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store all model fits
save(list = 'all.fit', file = 'MAYO_CBE_TCX_allfits.RData')
AFIT_OBJ = File('MAYO_CBE_TCX_allfits.RData', 
                name = 'All model fits (BrainRegion.Diagnosis)', 
                parentId = CODE$properties$id)
annotations(AFIT_OBJ) = annotations(CODE)
annotations(AFIT_OBJ)$dataSubType = 'modelFits'
AFIT_OBJ = synStore(AFIT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates                                 |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r ECOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Offset (for gene length and gc content)    |  `r OFFSET_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Surrogate Variables                        |  `r SURR_OBJ$properties$id`      |
|  Residual Expression (for eQTL)             |  `r eEXP_OBJ$properties$id`      |
|  Residual Expression (for network analysis) |  `r nEXP_OBJ$properties$id`      |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |

### 2017-09-05 code/Rmd/MSSM_REPROCESSED.Rmd
Minor modifications and added parallelDuplicateCorrelation.R with lmer

  # # Calculate correlation values of random effects
  correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
                                             block = COVARIATES$individualIdentifier)
  # correlation = list(); correlation$cor = 0.4

>>>>

  # # Calculate correlation values of random effects
  # correlation = parallelDuplicateCorrelation(VOOM.GENE_EXPRESSION,
  #                                            block = COVARIATES$individualIdentifier)
  correlation = list(); correlation$cor = 0.457197
